A New Metric for the Analysis of Swarms using Potential Fields

### 2018-08-07 paper.tex
Almost completed updates. Structure to update.

\begin{abstract}
  % An important criterion in the assessment of swarm coordination algorithms is
  % the extent to which they cause unnecessary movement, i.e. movement which
  % occurs as a side effect of the algorithm rather than from the fulfilment of
  % the intended function of the swarm. 
  There are many metrics defined for the analysis of swarm coordination algorithms. These metrics are usually based upon the distances between agents, the distance between agents and a fixed point or the resultant vector that \textit{potential field} effects produce. This paper examines a distance-based metric that measures a swarms overall structure using inter-agent distances and introduces an new metric that identifies a swarm's state based upon the resultant magnitude of the vectors produced by the inter-agent interactions that create the agent distribution with a swarm's structure. The algorithms used to implement the swarming feature are based upon mathematically generated cohesion and repulsion vectors between an agent and its neighbours. In comparing and contrasting the two metrics, we find that the \textit{cohesion/repulsion} metric offers a number of advantages over the distance metric. In particular, the \textit{cohesion/repulsion} metric allows the identification of the essential characteristic of a swarm as `expanding', `stable', or `contracting'. These state cannot be identified using a distance based metric. Practical swarming applications where the new metric can be applied include area filling and reconnaissance.
\end{abstract}

@@

Many coordination algorithms exist to control the movement of agents within a swarm. The primary goal of these algorithms is to maintain a stable swarm structure. A stable structure is one where the agents in the swarm ``act'' as a single entity. The main focus of this paper is the use of a metric that is based upon any coordination algorithms that use \textit{potential fields} to produce a swarming effect. To facilitate this analysis a basic swarm model is introduced and then the two metrics are applied to this model to present a comparison of the information provided by a \textit{distance-based} and the new \textit{cohesion/repulsion-based} metric. The analysis includes the identification of a swarm's ``state''.

@@

\section{Residual internal movement (Jitter)}\label{Section:Jitter}
Due to the dynamic nature of a swarm, maintaining optimum internal movement as in~Figure \ref{methods:StabilityNullVector} is highly unlikely. The agent pairs may fluctuate between the 4 states~(Fig.~\ref{methods:Stability1}, \ref{methods:Stability2}, \ref{methods:Stability3} and \ref{methods:Stability4}); These transitions are \textit{jitter}. The degree to which this variation occurs can be measured using either the change in distance between the agents, or the change in the magnitude $P(b)$~(Equation~\ref{eq:CohesionRepulsion}) between the agents. Jitter arises as motion to maintain the structure of a swarm. A swarm coordination algorithm that reduces jitter is generally desirable.

@@

Figure~\ref{methods:Stability1} shows two agents within each other's cohesion fields but sufficiently distant to be outside of the repulsion fields. In this case $|k_cv_c| > 0$ and $|k_rv_r| = 0$: the result is the agent's resultant magnitude $v(b)$ causes the agent to move towards is neighbour $b'$. Likewise the neighbour's resultant magnitude $v(b')$ will cause it to move towards $b$. Table~\ref{tab:SampleReplusion0} shows the repulsion magnitude with a value of 0. The only influence on the agent pairs are cohesive vectors.

@@

\section{Distance metric}\label{Section:DistanceDynamics}
The distance metric considers a swarm in terms of how the agents are physically distributed: i.e. only the inter-agent distances and their standard deviation are considered. The standard deviation indicates the extent to which the swarm is out of balance and will move to re-balance itself. If the standard deviation is zero then all the agents are evenly spaced.

@@

The distance-based metric for the internal distribution of the agents is the pair consisting of $\mu_d(S)$, $\sigma_d(S)$. This can be written informally as:

@@

% The \textit{`cohesion/repulsion metric'} is therefore the scalar value of the \textit{`agent resultant magnitude'} which is altered based upon the difference between the repulsion and cohesion magnitudes~(\ref{eq:CohesionRepulsion}).
This paper defines the cohesion/repulsion metric scalar for an agent as the length of $v(b)$  as defined in Equation~\ref{eq:InterAgentMovement2}: that is, the weighted effects of cohesion and repulsion by neighbours of the agent. This number carries a negative sign if repulsion dominates and a positive sign if cohesion dominates and a positive sign if cohesion donates. This is summarised in Equation~\ref{eq:CohesionRepulsion}.

@@

The detection of a swarm's perimeter~\cite{MD:09, MJ:08, ZAPS:07, JG:13}
identifies a subset of a swarm which can be used for coordination. 
Perimeter detection is widely used in swarm coordination~\cite{ZFG:13, AKK:08,
APZDAMC:09, AZDPS:11}. A possible adaptation of this area of research is to
investigate the reduction of the field effects (cohesion and repulsion) of
agents at the perimeter, so increasing agent density in this region. Increasing
the density would allow increased coverage by the swarm at its perimeter. This
could improve the granularity of analysis, e.g. when using sensor arrays at
thr frontier of the swarm.
However, this approach leads to swarms of agents having heterogeneous field
effects, as shown in Fig.~\ref{additional:fieldsWork}. Clearly, the
distance-based metric is not useful in the analysis of such swarms. We believe
that the cohesion/repulsion metric introduced in this paper can identify a
swarm's structure even when the swarm is heterogeneous, since the metric is
based directly upon inter-agent field effects rather than the physical
distribution of agents. This will also be the subject of forthcoming work.

>>>>

\begin{abstract}
  % An important criterion in the assessment of swarm coordination algorithms is
  % the extent to which they cause unnecessary movement, i.e. movement which
  % occurs as a side effect of the algorithm rather than from the fulfilment of
  % the intended function of the swarm. 
  There are many metrics defined for the analysis of swarm coordination algorithms. These metrics are usually based upon the distances between agents, the distance between agents and a fixed point or the resultant vector that \textit{potential field} effects produce. This paper examines a distance-based metric that measures a swarms overall structure using inter-agent distances and introduces an new metric that identifies a swarm's state based upon the resultant magnitude of the vectors produced by the inter-agent interactions that create the agent distribution within a swarm's structure. The algorithms used to implement the swarming feature are based upon mathematically generated cohesion and repulsion vectors between an agent and its neighbours. In comparing and contrasting the two metrics, we find that the \textit{cohesion/repulsion} metric offers a number of advantages over the distance metric. In particular, the \textit{cohesion/repulsion} metric allows the identification of the essential characteristic of a swarm as `expanding', `stable', or `contracting'. These state cannot be identified using a distance based metric. Practical swarming applications where the new metric can be applied include area filling and reconnaissance.
\end{abstract}

@@

Many coordination algorithms exist to control the movement of agents within a swarm. The primary goal of these algorithms is to maintain a stable swarm structure. A stable structure is one where the agents in the swarm ``act'' as a single entity. The main focus of this paper is the use of a metric that can be applied to any coordination algorithms that use \textit{potential fields} to produce a swarming effect. To facilitate this analysis a basic swarm model is introduced and then the two metrics are applied to this model to present a comparison of the information provided by a \textit{distance-based} and the new \textit{cohesion/repulsion-based} metric. The analysis includes the identification of a swarm's ``state''. The metric is then applied to a flood filling scenario to highlight a possible use of the new metric.

@@

% \section{Residual internal movement (Jitter)}\label{Section:Jitter}
% Due to the dynamic nature of a swarm, maintaining optimum internal movement as in~Figure \ref{methods:StabilityNullVector} is highly unlikely. The agent pairs may fluctuate between the 4 states~(Fig.~\ref{methods:Stability1}, \ref{methods:Stability2}, \ref{methods:Stability3} and \ref{methods:Stability4}); These transitions are \textit{jitter}. The degree to which this variation occurs can be measured using either the change in distance between the agents, or the change in the magnitude $P(b)$~(Equation~\ref{eq:CohesionRepulsion}) between the agents. Jitter arises as motion to maintain the structure of a swarm. A swarm coordination algorithm that reduces jitter is generally desirable.

@@

Figure~\ref{methods:Stability1} shows two agents within each other's cohesion fields but sufficiently distant to be outside of the repulsion fields. In this case $|k_cv_c| > 0$ and $|k_rv_r| = 0$: the result is the agent's resultant magnitude $v(b)$ causes the agent to move towards its neighbour $b'$. Likewise the neighbour's resultant magnitude $v(b')$ will cause it to move towards $b$. Table~\ref{tab:SampleReplusion0} shows the repulsion magnitude with a value of 0. The only influence on the agent pair are cohesive vectors.

@@

\section{Inter-Agent Distance metric}\label{Section:DistanceDynamics}
The inter-agent distance metric considers a swarm in terms of how the agents are physically distributed: i.e. only the inter-agent distances and their standard deviation are considered. The standard deviation indicates the extent to which the swarm is out of balance and therefore the degree of movement required to re-balance the swarm structure. If the standard deviation is zero then all the agents are evenly spaced.

@@

The distance-based metric for the internal distribution of the agents is therefore $\mu_d(S)$ and $\sigma_d(S)$. This can be written informally as:

@@

% The \textit{`cohesion/repulsion metric'} is therefore the scalar value of the \textit{`agent resultant magnitude'} which is altered based upon the difference between the repulsion and cohesion magnitudes~(\ref{eq:CohesionRepulsion}).
This paper defines the cohesion/repulsion metric scalar for an agent as the length of $v(b)$  as defined in Equation~\ref{eq:InterAgentMovement2}: that is, the weighted effects of cohesion and repulsion by neighbours of the agent. This number carries a negative sign if repulsion dominates and a positive sign if cohesion dominates. This is summarised in Equation~\ref{eq:CohesionRepulsion}.

@@

The detection of a swarm's perimeter~\cite{MD:09, MJ:08, ZAPS:07, JG:13}
identifies a subset of a swarm which can be used for coordination. 
Perimeter detection is widely used in swarm coordination~\cite{ZFG:13, AKK:08,
APZDAMC:09, AZDPS:11}. A possible adaptation of this area of research is to
investigate the reduction of the field effects (cohesion and repulsion) of
agents at the perimeter, so increasing agent density in this region. Increasing
the density would allow increased coverage by the swarm at its perimeter. This
could improve the granularity of analysis, e.g. when using sensor arrays at
the frontier of the swarm.
However, this approach leads to swarms of agents having heterogeneous field
effects, as shown in Fig.~\ref{additional:fieldsWork}. Clearly, the
distance-based metric is not useful in the analysis of such swarms. We believe
that the cohesion/repulsion metric introduced in this paper can identify a
swarm's structure even when the swarm is heterogeneous, since the metric is
based directly upon inter-agent field effects rather than the physical
distribution of agents. This will also be the subject of forthcoming work.
